---
title: "9-random-forest"
output: html_document
---

```{r}
library(tidyverse)
library(plotly)
library(doParallel)
library(foreach)
library(caret)
library(rpart)
library(vip)
library(pdp)
```

-   Optimal performance is often found by bagging 50–500 trees. Data sets that have a few strong predictors typically require less trees; whereas data sets with lots of noise or multiple strong predictors may need more. Using too many trees will not lead to overﬁtting. However, it’s important to realize that since multiple models are being run, the more iterations you perform the more computational and time requirements you will have. As these demands increase, performing k-fold CV can become computationally burdensome.

-   A beneﬁt to creating ensembles via bagging, which is based on resampling with replacement, is that it can provide its own internal estimate of predictive performance with the out-of-bag (OOB) sample. The OOB sample can be used to test predictive performance and the results usually compare well compared to k-fold CV assuming your data set is suﬃciently large (say 𝑛 ≥ 1, 000). Consequently, as your data sets become larger and your bagging iterations increase, it is common to use the OOB error estimate as a proxy for predictive performance.

```{r}
ames <- AmesHousing::make_ames()

ames$Date <- as.Date(paste(ames$Year_Sold, ames$Mo_Sold, "01", sep = "-"))

set.seed(40)  # for reproducibility
split  <- rsample::initial_split(ames, prop = 0.7, strata = "Sale_Price")
ames_train  <- rsample::training(split)
ames_test   <- rsample::testing(split)
```

```{r}
set.seed(40)

ames_bag1 <- train(
  Sale_Price ~ .,
  data = ames_train,
  method = "treebag",
  nbagg = 100,
  coob = TRUE,
  control = rpart.control(minsplit = 2, cp = 0)
)

ames_bag1
```

```{r}
# using ranger to do the same as above.  Will allow for bagging under 10 trees
# and is much faster!
ntree <- seq(1, 200, by = 2)
# create empty vector to store OOB RMSE values
rmse <- vector(mode = "numeric", length = length(ntree))

for (i in seq_along(ntree)) {
  # reproducibility
  set.seed(123)
  # perform bagged model
  model <- ranger::ranger(
  formula = Sale_Price ~ .,
  data    = ames_train,
  num.trees = ntree[i],
  mtry = ncol(ames_train) - 1,
  min.node.size = 1
)
  # get OOB error
  rmse[i] <- sqrt(model$prediction.error)
}

bagging_errors <- data.frame(ntree, rmse)

ggplot(bagging_errors, aes(ntree, rmse)) +
  geom_line() +
  geom_hline(yintercept = 41019, lty = "dashed", color = "grey50") +
  annotate("text", x = 100, y = 41385, label = "Best individual pruned tree", 
           vjust = 0, hjust = 0, color = "grey50") +
  annotate("text", x = 100, y = 26750, label = "Bagged trees", vjust = 0, hjust = 0) +
  ylab("RMSE") +
  xlab("Number of trees")
```

```{r}
ames_bag2 <- train(
  Sale_Price ~ .,
  data = ames_train,
  method = "treebag",
  trControl = trainControl(method = "cv", number = 10),
  nbagg = 200,
  control = rpart.control(minsplit = 2, cp = 0)
)

ames_bag2
```

```{r}
# Create a parallel socket cluster
cl <- makeCluster(8) # use 8 workers
registerDoParallel(cl) # register the parallel backend
```

```{r}
# Fit trees in parallel and compute predictions on the test set
predictions <- foreach(
  icount(160),
  .packages = "rpart",
  .combine = cbind
  ) %dopar% {
    # bootstrap copy of training data
    index <- sample(nrow(ames_train), replace = TRUE)
    ames_train_boot <- ames_train[index, ]
    
    # fit tree to bootstrap copy
    bagged_tree <- rpart(
      Sale_Price ~ .,
      control = rpart.control(minsplit = 2, cp = 0),
      data = ames_train_boot
    )
    
    predict(bagged_tree, newdata = ames_test)
}
predictions[1:5, 1:7]
```

```{r}
predictions %>%
  as.data.frame() %>%
  mutate(
    observation = 1:n(),
    actual = ames_test$Sale_Price
  ) %>%
  tidyr::gather(tree, predicted, -c(observation, actual)) %>%
  group_by(observation) %>%
  mutate(tree = stringr::str_extract(tree, '\\d+') %>% as.numeric()) %>%
  ungroup() %>%
  arrange(observation, tree) %>%
  group_by(observation) %>%
  mutate(avg_prediction = cummean(predicted)) %>%
  group_by(tree) %>%
  summarize(RMSE = RMSE(avg_prediction, actual)) %>%
  ggplot(aes(tree, RMSE)) +
  geom_line() +
  xlab('Number of trees')
```

```{r}
# Shutdown parallel cluster
stopCluster(cl)
```

```{r}
vip(ames_bag2, num_features = 40, bar = FALSE)
```

```{r}
# Construct partial dependence plots
p1 <- pdp::partial(
  ames_bag2,
  pred.var = 'Lot_Area',
  grid.resolution = 20
) %>% autoplot()

p2 <- pdp::partial(
  ames_bag2,
  pred.var = "Lot_Frontage",
  grid.resolution = 20
) %>% autoplot()
```

```{r}
gridExtra::grid.arrange(p1, p2, nrow = 1)
```

```{r}

```
