---
title: "project"
output: html_document
---

```{r}
library(tidyverse)
library(tidyr)
library(h2o)
library(TSstudio)
library(plotly)
library(zoo)
library(xts)
library(Quandl)
library(forecast)
library(readr)
library(naniar)
library(corrplot)
library(timetk)
library(tseries)
library(lmtest)
library(dtw)
library(reshape2)
library(gganimate)
library(WaveletComp)
library(tuneR)
library(seewave)
```

#### Description of Dataset Columns

-   psr: personal saving rate

-   unrate: unemployment rate

-   m2: m2

-   dspic: real disposable personal income

-   pce: personal consumption expenditure

-   reer: real broad effective exchange rate

-   ir: market yield on US securities at 10-year constant maturity

-   ffer: federal funds effective rate

-   tcs: total construction spending

-   indpro:

-   cci:

```{r}
macro_monthly <- read_csv("~/Work/Data Science/Learning/Datasets/macro-monthly/macro_monthly.csv")
head(macro_monthly)
```

```{r}
names(macro_monthly)
```

```{r}
macro_monthly <- macro_monthly %>%
  # mutate(across(where(is.double), as.integer))
  mutate(across(-1, as.numeric))  # Excludes the first column

head(macro_monthly)
```

```{r}
str(macro_monthly)
```

```{r}
ts_plot(macro_monthly)
```

```{r}
colSums(is.na(macro_monthly))
```

```{r}
vis_miss(macro_monthly)
```

```{r}
macro_monthly <- macro_monthly %>%
  tidyr::fill(c(reer, tcs), .direction = "downup")  # Forward fill (Down), Backward fill (Up)
```

```{r}
colSums(is.na(macro_monthly))
```

```{r}
macro_monthly_cor <- cor(macro_monthly[,-1])
macro_monthly_cor
```

```{r}
corrplot(macro_monthly_cor)
```

-   From the correlation plot of the correlation matrix, columns *unrate and psr* are correlated with a rho=0.5. Columns *m2, dspic* and *pce each* having a strong correlation of more than 0.9. ir and ffer with rho=0.93. indpro and ccpi with rho=0.85

```{r}
macro_monthly_ts <- ts(macro_monthly, frequency = 1)
macro_monthly_ts
```

```{r}
unrate_psr <- macro_monthly %>% select(DATE, unrate, psr)
m2_dspic_pce <- macro_monthly %>% select(DATE, m2, dspic, pce)
ir_ffer <- macro_monthly %>% select(DATE, ir, ffer)
indpro_ccpi <- macro_monthly %>% select(DATE, indpro, ccpi)
m2_dspic_pce_tcs_indpro_ccpi <- macro_monthly %>% select(DATE, m2, dspic, pce, tcs, indpro, ccpi)
dspic_pce_tcs_indpro_ccpi <- macro_monthly %>% select(DATE, dspic, pce, tcs, indpro, ccpi)

head(unrate_psr)
```

## Time Series Structure

```{r}
unrate_psr_xts <- as.xts(unrate_psr)
m2_dspic_pce_xts <- as.xts(m2_dspic_pce)
ir_ffer_xts <- as.xts(ir_ffer)
indpro_ccpi_xts <- as.xts(indpro_ccpi)
m2_dspic_pce_tcs_indpro_ccpi_xts <- as.xts(m2_dspic_pce_tcs_indpro_ccpi)
dspic_pce_tcs_indpro_ccpi_xts <- as.xts(dspic_pce_tcs_indpro_ccpi)

par(mfrow=c(2,2))
ts_plot(unrate_psr_xts, Xgrid = TRUE, Ygrid = TRUE)
ts_plot(m2_dspic_pce_xts, Xgrid = TRUE, Ygrid = TRUE)
ts_plot(ir_ffer_xts, Xgrid = TRUE, Ygrid = TRUE)
ts_plot(indpro_ccpi_xts, Xgrid = TRUE, Ygrid = TRUE)
# ts_plot(m2_dspic_pce_tcs_indpro_ccpi_xts)
# ts_plot(dspic_pce_tcs_indpro_ccpi_xts)
```

```{r}
names(macro_monthly)
```

```{r}
unrate <- macro_monthly %>% select(DATE, unrate)
psr <- macro_monthly %>% select(DATE, psr)
m2 <- macro_monthly %>% select(DATE, m2)
dspic <- macro_monthly %>% select(DATE, dspic)
pce <- macro_monthly %>% select(DATE, pce)
ir <- macro_monthly %>% select(DATE, ir)
ffer <- macro_monthly %>% select(DATE, ffer)
indpro <- macro_monthly %>% select(DATE, indpro)
ccpi <- macro_monthly %>% select(DATE, ccpi)

unrate_xts <- as.xts(unrate)
psr_xts <- as.xts(psr)
```

```{r}
head(unrate_xts)
```

```{r}
unrate$DATE <- as.Date(unrate$DATE)
psr$DATE <- as.Date(psr$DATE)
m2$DATE <- as.Date(m2$DATE)
dspic$DATE <- as.Date(dspic$DATE)
pce$DATE <- as.Date(pce$DATE)
ir$DATE <- as.Date(ir$DATE)
ffer$DATE <- as.Date(ffer$DATE)
indpro$DATE <- as.Date(indpro$DATE)
ccpi$DATE <- as.Date(ccpi$DATE)

head(unrate)
```

### 1. Decomposition

```{r}
unrate_ts <- tk_ts(unrate, 
                   start = c(lubridate::year(min(macro_monthly$DATE)), 
                             lubridate::month(min(macro_monthly$DATE))), 
                   frequency = 12)
psr_ts <- tk_ts(psr, 
                start = c(lubridate::year(min(macro_monthly$DATE)), 
                             lubridate::month(min(macro_monthly$DATE))), 
                frequency = 12)
m2_ts <- tk_ts(m2, 
               start = c(lubridate::year(min(macro_monthly$DATE)), 
                             lubridate::month(min(macro_monthly$DATE))), 
               frequency = 12)
dspic_ts <- tk_ts(dspic, 
                  start = c(lubridate::year(min(macro_monthly$DATE)), 
                             lubridate::month(min(macro_monthly$DATE))), 
                  frequency = 12)
pce_ts <- tk_ts(pce, 
                start = c(lubridate::year(min(macro_monthly$DATE)), 
                             lubridate::month(min(macro_monthly$DATE))), 
                frequency = 12)
ir_ts <- tk_ts(ir, 
               start = c(lubridate::year(min(macro_monthly$DATE)), 
                             lubridate::month(min(macro_monthly$DATE))), 
               frequency = 12)
ffer_ts <- tk_ts(ffer, 
                 start = c(lubridate::year(min(macro_monthly$DATE)), 
                             lubridate::month(min(macro_monthly$DATE))), 
                 frequency = 12)
indpro_ts <- tk_ts(indpro, 
                   start = c(lubridate::year(min(macro_monthly$DATE)), 
                             lubridate::month(min(macro_monthly$DATE))), 
                   frequency = 12)
ccpi_ts <- tk_ts(ccpi, 
                 start = c(lubridate::year(min(macro_monthly$DATE)), 
                             lubridate::month(min(macro_monthly$DATE))), 
                 frequency = 12)

ts_decompose(unrate_ts)
ts_decompose(psr_ts)
ts_decompose(m2_ts)
ts_decompose(dspic_ts)
ts_decompose(pce_ts)
ts_decompose(ir_ts)
ts_decompose(ffer_ts)
ts_decompose(indpro_ts)
ts_decompose(ccpi_ts)
```

```{r}
unrate_detrend <- unrate_ts - decompose(unrate_ts)$trend
psr_detrend <- psr_ts - decompose(psr_ts)$trend
m2_detrend <- m2_ts - decompose(m2_ts)$trend
dspic_detrend <- dspic_ts - decompose(dspic_ts)$trend
pce_detrend <- pce_ts - decompose(pce_ts)$trend
ir_detrend <- ir_ts - decompose(ir_ts)$trend
ffer_detrend <- ffer_ts - decompose(ffer_ts)$trend
indpro_detrend <- indpro_ts - decompose(indpro_ts)$trend
ccpi_detrend <- ccpi_ts - decompose(ccpi_ts)$trend

ts_seasonal(unrate_detrend)
ts_seasonal(psr_detrend)
ts_seasonal(m2_detrend)
ts_seasonal(dspic_detrend)
ts_seasonal(pce_detrend)
ts_seasonal(ir_detrend)
ts_seasonal(ffer_detrend)
ts_seasonal(indpro_detrend)
ts_seasonal(ccpi_detrend)
```

```{r}
simple_ma <- function(ts.obj, order) {
  lags <- order - 1
  lags <- make_lags(ts.obj = ts.obj, lags = 1)
  mean <- ts_mean(lags)
  un <- ts.union(ts.obj, mean)
  colnames(un) <- c("Original", "Transformed")
  
  return(un)
}
```

```{r}
simple_ma4_unrate <- simple_ma(unrate_ts, order = 4)
simple_ma4_psr <- simple_ma(psr_ts, order = 4)
simple_ma4_m2 <- simple_ma(m2_ts, order = 4)
simple_ma4_pce <- simple_ma(pce_ts, order = 4)
simple_ma4_dspic <- simple_ma(dspic_ts, order = 4)
simple_ma4_ir <- simple_ma(ir_ts, order = 4)
simple_ma4_ffer <- simple_ma(ffer_ts, order = 4)
simple_ma4_indpro <- simple_ma(indpro_ts, order = 4)
simple_ma4_ccpi <- simple_ma(ccpi_ts, order = 4)
```

```{r}
ts_plot(simple_ma4_unrate, type = "multiple", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(simple_ma4_psr, type = "multiple", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(simple_ma4_m2, type = "multiple", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(simple_ma4_pce, type = "multiple", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(simple_ma4_dspic, type = "multiple", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(simple_ma4_ir, type = "multiple", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(simple_ma4_ffer, type = "multiple", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(simple_ma4_indpro, type = "multiple", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(simple_ma4_ccpi, type = "multiple", Xgrid = TRUE, Ygrid = TRUE)
```

```{r}
ts_info(psr_ts)
```

```{r}
psr_cyclic <- window(psr_ts, start = c(2000, 1))
unrate_cyclic <- window(psr_ts, start = c(2000, 1))
m2_cyclic <- window(psr_ts, start = c(2000, 1))
pce_cyclic <- window(psr_ts, start = c(2000, 1))
dspic_cyclic <- window(psr_ts, start = c(2000, 1))
ir_cyclic <- window(psr_ts, start = c(2000, 1))
ffer_cyclic <- window(psr_ts, start = c(2000, 1))
indpro_cyclic <- window(psr_ts, start = c(2000, 1))
ccpi_cyclic <- window(psr_ts, start = c(2000, 1))

ts_plot(unrate_cyclic, type = "single", title = "US Monthly Unemployment Rate",
        Ytitle = "Unemployment Rate (%)",
        Xtitle = "Year", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(psr_cyclic, type = "single", title = "US Monthly Personal Saving Rate",
        Ytitle = "Personal Saving Rate (%)",
        Xtitle = "Year", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(m2_cyclic, type = "single", title = "US monthly m2",
        Ytitle = "m2",
        Xtitle = "Year", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(pce_cyclic, type = "single", title = "US Monthly Personal Consumption Expenditure",
        Ytitle = "Personal Consumption Expenditure",
        Xtitle = "Year", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(dspic_cyclic, type = "single", title = "US Monthly DIsposable Personal Income",
        Ytitle = "DIsposable Personal Income",
        Xtitle = "Year", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(ir_cyclic, type = "single", title = "US Monthly Market Yield on Securities",
        Ytitle = "Market Yield on Securities",
        Xtitle = "Year", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(ffer_cyclic, type = "single", title = "US Monthly Federal Funds Effective Rate",
        Ytitle = "Federal Funds Effective Rate (%)",
        Xtitle = "Year", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(indpro_cyclic, type = "single", title = "US Monthly ",
        Ytitle = "INDPRO",
        Xtitle = "Year", Xgrid = TRUE, Ygrid = TRUE)
ts_plot(ccpi_cyclic, type = "single", title = "US Monthly",
        Ytitle = "CCPI",
        Xtitle = "Year", Xgrid = TRUE, Ygrid = TRUE)
```

```{r}
ts_heatmap(unrate_ts, title = "Heatmap - the US Unemployment Rate")
ts_heatmap(psr_ts, title = "Heatmap - the US Personal Saving Rate")
ts_heatmap(m2_ts, title = "Heatmap - the US m2")
ts_heatmap(pce_ts, title = "Heatmap - the US Personal Consumption Expenditure")
ts_heatmap(dspic_ts, title = "Heatmap - the US DIsposable Personal Income")
ts_heatmap(ir_ts, title = "Heatmap - the US Market Yield on Securities")
ts_heatmap(ffer_ts, title = "Heatmap - the US Federal Funds Effective Rate")
ts_heatmap(indpro_ts, title = "Heatmap - the US ")
ts_heatmap(ccpi_ts, title = "Heatmap - the US ")
```

-   The columns seem to be cyclic rather than seasonal due to the bright vertical lines in each heatmap.

```{r}
x <- lapply(1:36, function(i) {
  p <- Box.test(unrate_ts, lag = i, type = "Ljung-Box")
  output <- data.frame(lag = i, p_value = p$p.value)
  return(output) 
}) %>% bind_rows

head(x)
```

```{r}
unrate_df <- data.frame(year = floor(time(unrate_ts)), month = cycle(unrate_ts), unrate = as.numeric(unrate_ts))

# Setting the month abbreviation and transforming it to a factor
unrate_df$month <- factor(month.abb[unrate_df$month], levels = month.abb)

head(unrate_df)
```

```{r}
unrate_summary <- unrate_df %>%
  group_by(month) %>%
  summarise(mean = mean(unrate), sd = sd(unrate))

unrate_summary
```

```{r}
plot_ly(data = unrate_summary,
        x = ~ month,
        y = ~ mean,
        type = "bar",
        name = "Mean"
        ) %>%
  layout(title = "The US Monthly Avearge Unemployment Rate",
         yaxis = list(title = "Mean", range = c(4, 7)))
```

```{r}
unrate_df <- data.frame(time = index(unrate_xts), unrate = as.numeric(unrate_xts))

str(unrate_df)
```

```{r}
unrate_df$hour <- lubridate::hour(unrate_df$time)
unrate_df$weekday <- wday(unrate_df$time, label = TRUE, abbr = TRUE)
unrate_df$month <- factor(month.abb[lubridate::month(unrate_df$time)], levels = month.abb)

head(unrate_df)
```

```{r}
# Summarise the series by its hourly cycle
unrate_hourly <- unrate_df %>%
  group_by(hour) %>%
  summarise(mean = mean(unrate, na.rm = TRUE), sd = sd(unrate, na.rm = TRUE))

head(unrate_hourly)
```

```{r}
unrate_daily <- unrate_df %>%
  group_by(weekday) %>%
  summarise(mean = mean(unrate, na.rm = TRUE), sd = sd(unrate, na.rm = TRUE))

unrate_daily
```

```{r}
plot_ly(data = unrate_daily,
        x = ~ weekday,
        y = ~ mean,
        type = "bar"
       ) %>%
  layout(title = "The Daily Unemployment Rate by Weekday",
         yaxis = list(title = "Mean", range = c(4, 8)),
         xaxis = list(title = "Weekday"))
```

-   The average of the weekdays show that a matching average value in some of the days. This shows that our data is not seasonal

```{r}
ggplot(unrate_df, aes(x = unrate)) +
  geom_density(aes(fill = month)) +
  ggtitle("Uemployment Rate - Kernel Density by Month") +
  facet_grid(rows = vars(as.factor(month)))
```

```{r}
unrate_df$unrate_detrend <- unrate_df$unrate - decompose(unrate_ts)$trend

ggplot(unrate_df, aes(x = unrate_detrend)) +
  geom_density(aes(fill = month)) +
  ggtitle("Unemployment Rate - Kernel Density Estimates by Month") +
  facet_grid(rows = vars(as.factor(month)))
```

```{r}
unrate_df$weekday <- as.factor(unrate_df$weekday)

unrate_df %>%
  ggplot(aes(x = unrate)) +
    geom_density(aes(fill = as.factor(weekday))) +
    ggtitle("Unemployment Rate - Kernel Density Estimates by Day") +
    facet_grid(rows = vars(as.factor(weekday)))
```

```{r}
ggplot(unrate_df, aes(x = unrate_detrend)) +
  geom_density(aes(fill = weekday)) +
  ggtitle("Unemployment Rate - Kernel Density Estimates by Month") +
  facet_grid(rows = vars(as.factor(weekday)))
```

### Correlation Analysis

```{r}
ts_cor(unrate_ts, type = "both", lag.max = 60)
ts_cor(psr_ts, type = "both", lag.max = 60)
ts_cor(m2_ts, type = "both", lag.max = 60)
ts_cor(dspic_ts, type = "both", lag.max = 60)
ts_cor(pce_ts, type = "both", lag.max = 60)
ts_cor(ir_ts, type = "both", lag.max = 60)
ts_cor(ffer_ts, type = "both", lag.max = 60)
ts_cor(indpro_ts, type = "both", lag.max = 60)
ts_cor(ccpi_ts, type = "both", lag.max = 60)
```

```{r}
ts_lags(unrate_ts, lags = c(12, 24, 36, 48, 60, 72))
ts_lags(psr_ts, lags = c(12, 24, 36, 48, 60, 72))
ts_lags(m2_ts, lags = c(12, 24, 36, 48, 60, 72))
ts_lags(dspic_ts, lags = c(12, 24, 36, 48, 60, 72))
ts_lags(pce_ts, lags = c(12, 24, 36, 48, 60, 72))
ts_lags(ir_ts, lags = c(12, 24, 36, 48, 60, 72))
ts_lags(ffer_ts, lags = c(12, 24, 36, 48, 60, 72))
ts_lags(indpro_ts, lags = c(12, 24, 36, 48, 60, 72))
ts_lags(ccpi_ts, lags = c(12, 24, 36, 48, 60, 72))
```

## Causality Tests

```{r}
psr_cyclic <- window(psr_ts, start = c(2000, 1))
unrate_cyclic <- window(psr_ts, start = c(2000, 1))
m2_cyclic <- window(psr_ts, start = c(2000, 1))
pce_cyclic <- window(psr_ts, start = c(2000, 1))
dspic_cyclic <- window(psr_ts, start = c(2000, 1))
ir_cyclic <- window(psr_ts, start = c(2000, 1))
ffer_cyclic <- window(psr_ts, start = c(2000, 1))
indpro_cyclic <- window(psr_ts, start = c(2000, 1))
ccpi_cyclic <- window(psr_ts, start = c(2000, 1))
```

```{r}
head(psr_cyclic)
```

```{r}
psr_unrate_combined <- cbind(psr_cyclic, unrate_cyclic)

psr_unrate_combined <- as.data.frame(psr_unrate_combined)

head(psr_unrate_combined)
```

```{r}
plot_ly(x = time(psr_cyclic), y = ~ psr_cyclic, type = "scatter", 
        mode = "line", name = "Personal Savings Rate") %>%
  add_lines(x = time(unrate_cyclic), y = ~ unrate_cyclic, name = "Unemployment Rate", yaxis = "y2") %>%
  layout(title = "Total Monthly Personal Saving vs Unemployment Rate in the US",
         yaxis2 = list(overlaying = "y", side = "right", title = "Percentage UNRate", showgrid = TRUE),
         yaxis = list(title = "Percentage PSR", showgrid = TRUE),
         legend = list(orientation = 'h'),
         margin = list(l = 50, r = 50, b = 50, t = 50, pad = 2))
```

```{r}
ccf_plot(x = psr_ts[,1], y = unrate_ts[,1], lags = 0:12)
ccf_plot(x = pce_ts[,1], y = dspic_ts[,1], lags = 0:12)
ccf_plot(x = ir_ts[,1], y = ffer_ts[,1], lags = 0:12)
ccf_plot(x = indpro_ts[,1], y = ccpi_ts[,1], lags = 0:12)
```

```{r}
head(psr_ts)
```

```{r}
ccf(x = as.numeric(psr_ts), y = as.numeric(unrate_ts), lags = 0:12)
```

#### Granger Causality

-   To check if one time series can predict another:

```{r}
grangertest(psr_ts ~ unrate_ts, order = 2)  # Test if unrate_ts causes psr_ts
```

```{r}
grangertest(unrate_ts ~ psr_ts, order = 2)  # Test if psr_ts causes unrate_ts
```

-   If `p-value < 0.05`, the independent variable significantly influences the dependent variable.

```{r}
lags <- 1:12
p_values <- sapply(lags, function(lag) grangertest(psr_ts ~ unrate_ts, order = lag)$`Pr(>F)`[2])

df_gc <- data.frame(Lag = lags, P_Value = p_values)
ggplot(df_gc, aes(x = Lag, y = P_Value, fill = P_Value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0.05) +
  labs(title = "Granger Causality Heatmap", x = "Lag", y = "P-Value") +
  theme_minimal()
```

#### **Cointegration Test (Engle-Granger)**

If two time series are non-stationary but move together over time, they may be cointegrated.

```{r}
adf.test(psr_ts)  # Check stationarity
adf.test(unrate_ts)
```

```{r}
# If both are non-stationary, test for cointegration
residuals <- lm(psr_ts ~ unrate_ts)$residuals
adf.test(residuals)  # If p < 0.05, they are cointegrated
```

-   The two series are not cointegrated

#### **Dynamic Time Warping (DTW)**

-   If the time series are not aligned, DTW can compare them:

```{r}
alignment <- dtw(psr_ts, unrate_ts)
plot(alignment)
```

#### Spectral Analysis

-   To determine dominant frequencies in time series.

```{r}
spectrum(psr_ts, main = "Periodogram of psr_ts")
spectrum(unrate_ts, main = "Periodogram of unrate_ts")
```

```{r}
ccf_values <- ccf(psr_ts[,1], unrate_ts[,1], lag.max = 12, plot = FALSE)
lags <- ccf_values$lag
correlations <- ccf_values$acf

lags_cors_df <- data.frame(lag = lags, correlation = correlations)
ggplot(lags_cors_df, aes(x = lag, y = correlation, fill = correlation)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
  labs(title = "Cross-Correlation Heatmap", x = "Lag", y = "Correlation") +
  theme_minimal()
```

#### **Rolling Window Correlation (Animated!)** 📊

-   To track how correlation changes over time

```{r}
rolling_corr <- rollapply(cbind(psr_ts, unrate_ts), width = 12,
                          FUN = function(x) cor(x[,1],x[,2], use = "complete.obs"),
                          by.column = FALSE, align = "right")

df_rolling <- data.frame(time = time(rolling_corr), correlation = rolling_corr)
head(df_rolling)
```

```{r}
# change the type of time column to numeric in order to plot the time correlation
df_rolling$time <- as.numeric(as.character(df_rolling$time))
df_rolling
```

```{r}
ggplot(df_rolling, aes(x = time, y = correlation)) +
  geom_line(color = "blue", size = 1) +
  labs(title = "Rolling Correlation Over Time", y = "Correlation", x = "Time") +
  theme_minimal() +
  transition_reveal(time)
```

**🔥 Why?**

-   Shows **how correlation strength changes dynamically** over time.

-   Can detect **shifts in trends**.

#### **Wavelet Coherence (For Non-Stationary Data)** 🌊

-   If trends shift **over time and frequency**, use **wavelet coherence** analysis:

```{r}
psr_unrate_ts <- data.frame(time = 1:length(psr_ts), psr = psr_ts, unrate = unrate_ts)

wave <- analyze.coherency(psr_unrate_ts, my.pair = c("psr","unrate"), 
                          loess.span = 0, make.pval = TRUE, n.sim = 100)

wc.image(wave, n.levels = 250, legend.params = list(lab = "Wavelet Coherence"))
```

**🔥 Why?**

-   Detects **both time and frequency relationships**.

-   Finds **hidden correlations that change over time**.

```{r}

```

## FORECASTING

```{r}
partition <- ts_split(unrate_ts, sample.out = 12)

unrate_train <- partition$train
unrate_test <- partition$test
```

```{r}
auto_arima_mdl <- auto.arima(unrate_train)

summary(auto_arima_mdl)
```

```{r}
checkresiduals(auto_arima_mdl)
```

-   The Ljung-Box test p-value is higher tha 0.05, therfore, we reject the null hypothesis with a level of significance 0.01. Hence, there is no correlation between residual series and its lags are different from zero.

-   The best fitted arima model is ARIMA(0,1,0)

-   The ACF plot can provide additional support for that as well. This indicates the model did not capture all of the series patterns, and may want to tune the model parameters.

```{r}
model_forecast <- forecast(auto_arima_mdl, h = 12)

accuracy(model_forecast, unrate_test)
```

```{r}
test_forecast(actual = unrate_ts, forecast.obj = model_forecast, test = unrate_test)
```

## Forecasting Using Linear Regression

```{r}
unrate_df <- ts_to_prophet(unrate_ts)

head(unrate_df)
```

```{r}
unrate_df$trend <- 1:nrow(unrate_df)

head(unrate_df)
```

```{r}
unrate_df$seasonal <- factor(lubridate::month(unrate_df$ds, label = TRUE), ordered = FALSE)

head(unrate_df)
```

```{r}
h <- 12 # setting a testing partition length

unrate_train <- unrate_df[1:(nrow(unrate_df) - h), ]
unrate_test <- unrate_df[(nrow(unrate_df) - h + 1):nrow(unrate_df), ]
```

```{r}
lm_model_trend <- lm(y ~ trend, data = unrate_df)
summary(lm_model_trend)
```

```{r}
unrate_train$yhat <- predict(lm_model_trend, newdata = unrate_train)

unrate_test$yhat <- predict(lm_model_trend, newdata = unrate_test)
```

```{r}
lm_model_plot <- function(data, train, test, title = NULL) {
  p <- plot_ly(data = data,
               x = ~ ds, y = ~ y,
               type = "scatter",
               mode = "line",
               name = "Actual") %>%
    add_lines(x = ~ train$ds,
              y = ~ train$yhat,
              line = list(color = "red"),
              name = "Fitted") %>%
    add_lines(x = ~ test$ds,
              y = ~ test$yhat,
              line = list(color = "green", dash = "dot", width = 3),
              name = "Forecasted") %>%
    layout(title = title,
           xaxis = list(title = "Year"),
           yaxis = list(title = "Billion Cubic Feet"),
           legend = list(x = 0.05, y = 0.95))
  
  return(p)
}
```

```{r}
lm_model_plot(data = unrate_df, train = unrate_train, test = unrate_test,
              title = "Predicting Trend of Unemployment Rate")
```

```{r}
trend_mape <- c(mean(abs(unrate_train$y - unrate_train$yhat) / unrate_train$y),
                mean(abs(unrate_test$y - unrate_test$yhat) / unrate_test$y))
trend_mape
```

```{r}
lm_model_seasonal <- lm(y ~ seasonal, data = unrate_train)
summary(lm_model_seasonal)
```

```{r}
unrate_train$yhat <- predict(lm_model_seasonal, newdata = unrate_train)
unrate_test$yhat <- predict(lm_model_seasonal, newdata = unrate_test)

lm_model_plot(data = unrate_df, train = unrate_train, test = unrate_test, 
              title = "Predicting Seasonality of the Series")
```

```{r}
seasonal_mape <- c(mean(abs(gas_train$y - gas_train$yhat) / gas_train$y),
                mean(abs(gas_test$y - gas_test$yhat) / gas_test$y))
seasonal_mape
```

```{r}
lm_model_comb <- lm(y ~ seasonal + trend, data = unrate_train)
summary(lm_model_comb)
```

```{r}
unrate_train$yhat <- predict(model_comb, newdata = unrate_train)
unrate_test$yhat <- predict(model_comb, newdata = unrate_test)

lm_model_plot(data = unrate_df,
              train = unrate_train,
              test = unrate_test,
              title = "Predicting Trend and Seasonality Components")
```

```{r}
combined_mape <- c(mean(abs(gas_train$y - gas_train$yhat) / gas_train$y),
                mean(abs(gas_test$y - gas_test$yhat) / gas_test$y))
combined_mape
```

```{r}
unrate_split <- ts_split(unrate_ts, sample.out = 12)

train.ts <- unrate_split$train
test.ts <- unrate_split$test
```

```{r}
tslm_model <- tslm(unrate_ts ~ season  + (season^2) + trend + I(trend^2), data = unrate_df)
summary(tslm_model)
```

```{r}
anova(tslm_model)
```

```{r}
unrate_df <- unrate_df %>%
  mutate(
    weekday = wday(ds, label = TRUE),
    month = lubridate::month(ds, label = TRUE),
    lag365 = dplyr::lag(y, 365)
  ) %>%
  filter(!is.na(lag365)) %>%
  arrange(ds)

head(unrate_df)
```

```{r}
tslm_forecast <- forecast(tslm_model)
```

## Forecasting Using Exponential Smoothing

```{r}
simple_moving_average <- function(df, horizon, m, weight = NULL) {
  # Error handling
  if (horizon > nrow(df)) {
    stop("The length of the forecast horizon must be shorter than the length of the series")
  }
  
  if (m > nrow(df)) {
    stop("The length of the rolling windoe must be shorter than the length of the series.")
  }
  
  if (!is.null(weight)) {
    if (length(weight) != m) {
      stop("The weight argument is not aligned with the length of the rolling window.")
    } else if (sum(weight) != 1) {
      stop("The sum of the average weight is different")
    }
  }
  
  # Setting the average weights
  if (is.null(weight)) {
    weight <- rep(1 / m, m)
  }
  
  # Setting the dataframe
  names(df)[1] <- "date"
  # Set train and test sets according to forecast horizon
  df$type <- c(rep("train", nrow(df) - horizon), 
               rep("test", horizon))
  # Spreading the table by partition
  df1 <- df %>% spread(key = type, value = y)
  
  # create target variable
  df1$yhat <- df1$train
  
  # simple moving average
  for (i in (nrow(df) - horizon + 1):nrow(df1)) {
    r <- (i - m):(i - 1)
    df1$yhat[i] <- sum(df1$yhat[r] * weight)
  }
  
  # dropping from the yhat variable the actual values
  # that were used for the rolling window
  df1$yhat <- ifelse(is.na(df1$test), NA, df1$yhat)
  
  df1$y <- ifelse(is.na(df1$test), df1$train, df1$test)
  
  return(df1)
}
```

```{r}
head(unrate)
```

```{r}
unrate_df <- ts_to_prophet(unrate_ts)
head(unrate_df)
```

```{r}
typeof(unrate_df)
```

```{r}
unrate_fc_m1 <- simple_moving_average(unrate_df, horizon = 24, m = 1)
unrate_fc_m6 <- simple_moving_average(unrate_df, horizon = 24, m = 6)
unrate_fc_m12 <- simple_moving_average(unrate_df, horizon = 24, m = 12)
unrate_fc_m24 <- simple_moving_average(unrate_df, horizon = 24, m = 24)
unrate_fc_m36 <- simple_moving_average(unrate_df, horizon = 24, m = 36)
unrate_fc_m48 <- simple_moving_average(unrate_df, horizon = 24, m = 48)
unrate_fc_m60 <- simple_moving_average(unrate_df, horizon = 24, m = 60)
```

```{r}
names(unrate_fc_m1)
```

```{r}
plot_ly(data = unrate_df[450:nrow(unrate_df), ], x = ~ ds, y = ~ y, type = "scatter", 
        mode = "lines", name = "Actual") %>%
  add_lines(x = unrate_fc_m1$date, y = unrate_fc_m1$yhat, name = "Simple Moving Average - 1",
            line = list(dash = "dash")) %>%
  add_lines(x = unrate_fc_m6$date, y = unrate_fc_m6$yhat, name = "Simple Moving Average - 6",
            line = list(dash = "dash")) %>%
  add_lines(x = unrate_fc_m12$date, y = unrate_fc_m12$yhat, name = "Simple Moving Average - 12",
            line = list(dash = "dash")) %>%
  add_lines(x = unrate_fc_m24$date, y = unrate_fc_m24$yhat, name = "Simple Moving Average - 24",
            line = list(dash = "dash")) %>%
  add_lines(x = unrate_fc_m36$date, y = unrate_fc_m36$yhat, name = "Simple Moving Average - 36",
            line = list(dash = "dash")) %>%
  add_lines(x = unrate_fc_m48$date, y = unrate_fc_m48$yhat, name = "Simple Moving Average - 48",
            line = list(dash = "dash")) %>%
  add_lines(x = unrate_fc_m60$date, y = unrate_fc_m60$yhat, name = "Simple Moving Average - 60",
            line = list(dash = "dash")) %>%
  layout(title = "Forecasting Unemployment Rates",
         xaxis = list(title = ""),
         yaxis = list(title = "Percentage (%)"))
```

```{r}
dim(unrate_df)
```

```{r}
# Weighted moving average
unrate_fc_wma12 <- simple_moving_average(unrate_df, horizon = 24, m = 12, weight = c(1, rep(0, 11)))


```
